---
title: "Week_12_Notes"
format: html
---

# Helpers, Build Systems, Environments, Packages

```{r}
library(here)
library(socviz)
library(tidyverse)
```

## helpers

### reprex

We need a reprex!! smallest possible way to reproduce the problem

```{r}
library(reprex)
```

```{r}
library(tidyverse)

starwars |> 
  count(homeworld, species) |> 
  mutate(pct = n / sum(n) * 100) |> 
  arrange(desc(pct))
```

copy ALL code to clipboard, then run "reprex()" in the console (yes with nothing in it!)

it will show you how the code would run in a brand new R session

then you can paste! (if you wanna ask for help somewhere)

``` r
library(tidyverse)
#> Warning: package 'tidyverse' was built under R version 4.4.3
#> Warning: package 'ggplot2' was built under R version 4.4.3
#> Warning: package 'tibble' was built under R version 4.4.3
#> Warning: package 'purrr' was built under R version 4.4.3
#> Warning: package 'stringr' was built under R version 4.4.3

starwars |> 
  count(homeworld, species) |> 
  mutate(pct = n / sum(n) * 100) |> 
  arrange(desc(pct))
#> # A tibble: 57 × 4
#>    homeworld species      n   pct
#>    <chr>     <chr>    <int> <dbl>
#>  1 Tatooine  Human        8  9.20
#>  2 <NA>      Human        6  6.90
#>  3 Naboo     Human        5  5.75
#>  4 Alderaan  Human        3  3.45
#>  5 Naboo     Gungan       3  3.45
#>  6 <NA>      Droid        3  3.45
#>  7 Corellia  Human        2  2.30
#>  8 Coruscant Human        2  2.30
#>  9 Kamino    Kaminoan     2  2.30
#> 10 Kashyyyk  Wookiee      2  2.30
#> # ℹ 47 more rows
```

<sup>Created on 2025-11-17 with [reprex v2.1.1](https://reprex.tidyverse.org)</sup>

reprex has different output formats (r code, plaintext, etc.)

### usethis

automates setup processes for projects, git repositories, etc.

can turn a random file into a project with create_project()

this is how to turn an R project into a repository: use_git() followed by use_github()

### tables

tables hard :( tables mostly ugly :(

-   tinytable

    -   basic function is tt()

    -   can style with other stuff in the package

```{r}
mtcars |> 
  tinytable::tt()    # look how cutie!
```

-   gt

    -   mainly makes html, not pdf

    -   also gtsummary as a package

    -   basically default to tinytable

-   skimr

    -   making codebook for self use (NOT publication)

```{r}
organdata |> skimr::skim(where(is.numeric)) |> skimr::partition()
```

## when projects get big

break them up! everything can be modular

creating coordinating scripts, functions, and packages

### Make

when your qmd relies on R scripts, Makefile manages these dependencies

everything is a DAG bc of course :(

but this is evil

### Targets

basically Make but for R yay!

super overkill for basic projects

### renv

freeze a file/project so that package/R updates don't break everything

aka uses old versions of things even if you have downloaded the new versions
