---
title: "Week_07_Notes"
format: html
---

# Ingesting Data

## (with haven and readr)

```{r}
library(tidyverse)
library(here)
library(haven)
library(socviz)
```

## csv files

usually our best friend!

but they are maybe actually evil bc sometimes data has commas in it and then everything is bad :(

tidyverse: read_csv() \[use this instead of read.csv()\]

read_csv2 uses ; instead of , for separators bc europe uses , for . in numbers

## other crazy things

read_line(): reads in lines as a vector rather than as a dataframe

read_table(): for tables that just use random spaces to divide columns

## remote reading files

```{r}
organ_remote <- read_csv("http://kjhealy.co/organdonation.csv")

organ_remote
```

the function will tell you how it read in the columns—make sure to look at this!!!

use "spec(x)" to see them

## normalizing names and recoding

janitor::clean_names() : this names names all lowercase, removes spaces, etc

## finding problems

if it gives you warnings when reading something in, you can use "problems(x)" to see what went wrong

you can specify within read_csv() to tell it how to read in certain columns (yay!)

## removing missing values

janitor::remove_empty("rows") : remove rows where all columns are empty (there is an inverse for columns)

## foreign formats

aka booooo stata

read_stata() is in haven

stata often comes with dbl+lbl column types

can drop all labels with zap_labels() and zap_missing() but you don't really wanna do this

so instead you can convert the labels with mutate and across

## partial application of functions

purr package has partial() which does... something

making your own mini functions, maybe? simple little guys?

can make a bunch of them and put them in a file or package so you can use them a bunch

## more factors

forcats package!

fct_recode(), fct_relabel(), fct_relevel(), etc.
