---
title: "Week_10_Notes"
format: html
---

# APIs and Databases

## Iterating on the Census

```{r}
library(tidyverse)
library(tidycensus)
```

one specific county

```{r}
out <- get_acs(geography = "county",
               variables = "B19013_001",
               state = "NY",
               county = "New York",
               survey = "acs1",
               year = 2005)

out
```

all counties

```{r}
out2 <- get_acs(geography = "county",
               variables = "B19013_001",
               state = "NY",
               survey = "acs1",
               year = 2005)

out2
```

but if you omit year, you just get the most recent year.

but we want all the years! so we must map()

```{r}
df <- 2005:2019 |> 
  set_names() |> 
  map(\(x) get_acs(geography = "county",
               variables = "B19013_001",
               state = "NY",
               survey = "acs1",
               year = x)) |> 
  list_rbind(names_to = "year") |> 
  mutate(year = as.integer(year))

df
```

## NYC Citibike API

```{r}
gbfsurl <- "https://gbfs.citibikenyc.com/gbfs/2.3/gbfs.json"
feeds <- jsonlite::fromJSON(gbfsurl)
str(feeds)
```

```{r}
feeds_df <- as_tibble(feeds)
feeds_df
```

look just at english data

```{r}
feeds_df |> 
  slice(1) |> 
  unnest(data) |>
  unnest(data)
```

this helps us find the specific url we need to get the specific data that we want

# Databases

## "Big" Data

it's just data but bigger (there is more of it)

data that is larger than your laptop's memory and makes it give up on life :(

but most of the time we don't want *all* the data; we are only interested in a small subset (certain rows or columns)

## Database Layouts

a database is not one giant table; instead it is like a list of tables connected through keys

this allows for clear and efficient storage of different levels of observation

## Database Normalization

taking your awful megatable and storing it in a way that doesn't suck

minimizing redundancy and the potential for error

(like IPEDS data where there is the ID in all the tables so you can link them)???

## Talking to Databases

usually done through an API or another DBI (DataBase Interface)

but let's do it manually for fun

open a connection

```{r}
library(DBI)

con <- dbConnect(duckdb::duckdb(), path = ":memory:")
```

move the data into the database server

R doesn't see it anymore, only the database does

```{r}
copy_to(
  dest = con,
  df = nycflights13::flights,
  name = "flights",
  temporary = FALSE,
  indexes = list(
    c("year", "month", "day"),
    "carrier",
    "tailnum",
    "dest"
  )
)
```

make a tibble

except what it puts out only *looks* like a tibble

```{r}
flights_db <- tbl(con, "flights")

flights_db
```

you can do stuff with it though!

but it won't count rows it hates doing that

```{r}
f_db <- flights_db |> 
  select(year:day, dep_delay, arr_delay)
```

but it is ***not actually doing anything***. this is all happening within the database dreamland

you won't get actual results back until you ask for them

```{r}
flights <- f_db |> 
  collect()

flights
```

now it exists for realsies.

doing shit with multiple tables

```{r}
copy_to(
  dest = con,
  df = nycflights13::planes,
  name = "planes",
  temporary = FALSE,
  indexes = "tailnum"
)

dbListTables(con)
```

```{r}
planes_db <- tbl(con, "planes")
```

```{r}
left_join(
  flights_db,
  planes_db |> rename(year_built = year),
  by = "tailnum"
) |> 
  select(year, month, day, dep_time, arr_time, carrier, flight, 
         tailnum, year_built, type, model)
```

\^ still not done for realsies BUT that shit together now

**finishing up: close the connection**

```{r}
dbDisconnect(con)
```
