---
title: "Week_11_Notes"
format: html
---

# Parallel Computing

```{r}
library(here)
library(socviz)
library(tidyverse)
```

## pipelines

the data and each step is processed sequentially/serially

bottlenecks at the CPU as tasks get bigger

data runs from hard drive to RAM to CPU; the slowdown can happen at any step

```{r}
mtcars |> 
  group_split(cyl) |> 
  map(\(df) lm(mpg ~ wt + hp + gear, data = df)) |> 
  map(summary) |> 
  map_dbl("r.squared")
```

```{r}
mtcars |> 
  group_by(cyl) |> 
  nest() |> 
  mutate(model = map(data, \(df) lm(mpg ~ wt + hp + gear, data = df)),
         perf = map(model, broom::glance)) |> 
  unnest(perf)
```

## embarrassingly parallel

when you split into groups, what happens in each group is independent from the others

so they can happen in any order, or at the same time

and then the results can be combined at the end

\*\* there is some bookkeeping: have to remember where you sent each piece and when they come back. but! you are responsive for less as the parallel portion increases

## multicore computing

parallel computing used to mean a giant room full of computers

but now CPUs have multiple cores

### terms

-   socket: physical connection on your computer that houses the processor

-   core: part of the processor that performs the computation. most modern processors have independent cores that can do wholly independent work. most of the time, most cores are not being used

-   process: single instance of running a task or program (Chrome, R, etc.). A core can only run one process at a time. BUT cores are super super fast, so they can run many *threads* (like a juggler with many balls in the air)

-   cluster: collection of things capable of hosting cores. can be a single socket on a laptop or a giant computer room

```{r}
parallel::detectCores()
```

## doing the thing

we used to have to use future and furrr, but now we can use purrr! yay!

```{r}
library(mirai)
status()
```

making demons! don't make more demons than you have cores.

```{r}
daemons(10)

status()
```

do parallel computing with the function "in_parallel"

\*\* there was a slow function example in class but i did not type it fast enough

```{r}
tictoc::tic()

parallel_out <- 1:20 |> 
  map(in_parallel(\(x) sqrt(x),
                  sqrt = sqrt))

tictoc::toc()
```

if you use in_parallel without setting daemons, it will just run as if in_parallel wasn't there. so no need to double-write code!

you MUST explicitly pass through the names of any local functions or objects. for functions not in base r, you also must fully namespace

```{r}
my_data <- c(1, 2, 3)

map(1:3, in_parallel(\(x) mean(my_data))) # this will not work

map(1:3, in_parallel(\(x) mean(my_data), my_data = my_data))
```

## real example

ok real example not gonna be here rn but it is in the slides

# Scraping

i hate scraping.

websites can by static/server-side or dynamic/client-side. static pages are evil. dynamic pages are so much more evil.

the minute i am asked to scrape again is the minute i drop out.
