---
title: "Week_04_Notes"
format: html
---

```{r}
library(tidyverse)
```

## Things to know about r

### 1.  It is a calculator (can do arithmetic and logic)

== for equality, ! for negation, != for inequality, etc.

### 2.  Everything in r has a name

some names are forbidden! like "TRUE" "FALSE" "NA" "NULL" "for" "if" etc. 
keep lowercase for simplicity. 

### 3. You can create objects

assign things to a name to create an object
we love concatenating! "c()" 

### 4. You can do things to objects with functions

functions always have () after their name
functions take inputs, perform actions, and return outputs
you can use ? with the function name to access help documentation
functions can be nested but instead you should use the pipe |> 
functions (objects) are bundled in packages

### Vector types; Object classes

objects are made of one or more vectors
vectors can only have a single type; objects can have vectors of different types
data frames are technically just lists

## Making Graphs

```{r}
library(here)
```

```{r}
df <- read_csv(here("Week_02/mptc_week_02/data", "ef2023d.csv"))
```




